---
title: "Homework#2"
author: "BMC"
date: "3/23/2021"
output: html_document
---


                                  Challenge #1
                               ------------------
```{r}
f <-"https://raw.githubusercontent.com/difiore/ADA-datasets/master/IMDB-movies.csv"
d <- read.csv(f)
d
d2 <- tibble(d)
class(d2)
```

 -  Use a one-line statement to filter the dataset to include just movies from 1920 to 1979 and movies that are between 1 and 3 hours long (runtimeMinutes >= 60 and runtimeMinutes <= 180), and add a new column that codes the startYear into a new variable, decade (“20s,” “30s,” …“70s”). If you do this correctly, there should be 5651 movies remaining in the dataset.
 
 

```{r}
d3 <- filter(d2,startYear>= 1920 & startYear <=1979
            & runtimeMinutes >= 60 & runtimeMinutes <= 180)%>%
mutate(d2,Decade = as.character(floor(startYear/10)*10 -1900))
d3
```

Use {ggplot2} to plot histograms of the distribution of runtimeMinutes for each decade.
```{r}
ggplot(data = d3, aes(x = runtimeMinutes))+ geom_histogram()+ facet_wrap(~Decade)
```


 Use a one-line statement to calculate the population mean and population standard deviation in runtimeMinutes for each decade and save the results in a new dataframe, results.
 
```{r}
results <- cbind.data.frame(aggregate(d3$runtimeMinutes, list(d3$Decade),mean),aggregate(d3$runtimeMinutes, list(d3$Decade),sd))
results
```


Draw a single sample of 100 movies, without replacement, from each decade
```{r}
d4 <- group_by(d3, Decade)
d5 <- sample_n(d4,100, replace=FALSE)
d5
```

Calculate this single sample mean and sample standard deviation in runtimeMinutes

```{r}
Mean1 <- mean(d5$runtimeMinutes)
Mean1
SD1 <- sd(d5$runtimeMinutes)
SD1
SE <-  Mean1 / sqrt(600)
SE
```
Mean = 98.6
SD = 21.2
SE = 4.02

Estimate the SE around the population mean runtimeMinutes for each decade based on the standard deviation and sample size from these samples.
```{r}
SD <- aggregate(d3$runtimeMinutes, list(d3$Decade),sd)
SD       
```

```{r}
# since the length of each decade is 100 ... the sqrt of 100 is 10 so I will divide the SD of each decade by 10 to get each decades SE. 
sigma.TW <- SD[1,]
sigma.TW
SE2 <- 26.2 / sqrt(100)
SE2
SE3 <- 17.3 /10
SE3
SE4 <-  1.92
SE5 <- 1.92
SE6 <- 2.12
SE7 <- 1.8
...
```
SE of 20s = 2.62
SE of 30s = 1.73
SE of 40s = 1.92
SE of 50s = 1.92
SE of 60s = 2.12
SE of 70s = 1.80

Compare these estimates to the actual population mean runtimeMinutes for each decade and to the calculated SE in the population mean for samples of size 100 based on the population standard deviation for each decade.

- The SE of the whole sample is 4.02, which is greater than the SE of any decade. The whole sample consists of 600 observations and each decade only consits of 100 observations so it makes sense that the SE of the decades would be less than the SE of the whole, large population. The SE of the mean of runtime minutes in the 20s stands out as greater when compared to every other decade. 


Generate a sampling distribution of mean runtimeMinutes for each decade by [a] drawing 1000 samples of 100 movies from each decade and, for each sample, [b] calculating the mean runtimeMinutes and the standard deviation in runtimeMinutes. Then, calculate the MEAN and the STANDARD DEVIATION of the sampling distribution of sample means for each decade (the former should be a very good estimate of the population mean, while the latter is another estimate of the standard error in the population mean for a particular sample size) and plot a histogram of the sampling distribution. What shape does it have?


[a] Generate a sampling distribution of mean runtimeMinutes for each decade by drawing 1000 samples of 100 movies from each decade and [b] calculating the mean runtimeMinutes

```{r}
reps <- 1000
n <- 100
samp_dist_mean20 <-
  do(reps) * mean(~runtimeMinutes, data = sample_n(filter(d3, Decade=="20"), size = n, replace = FALSE))

samp_dist_mean30 <-
  do(reps) * mean(~runtimeMinutes, data = sample_n(filter(d3, Decade=="30"), size = n, replace = FALSE))

samp_dist_mean40 <-
  do(reps) * mean(~runtimeMinutes, data = sample_n(filter(d3, Decade=="40"), size = n, replace = FALSE))

samp_dist_mean50 <-
  do(reps) * mean(~runtimeMinutes, data = sample_n(filter(d3, Decade=="50"), size = n, replace = FALSE))

samp_dist_mean60 <-
  do(reps) * mean(~runtimeMinutes, data = sample_n(filter(d3, Decade=="60"), size = n, replace = FALSE))

samp_dist_mean70 <-
  do(reps) * mean(~runtimeMinutes, data = sample_n(filter(d3, Decade=="70"), size = n, replace = FALSE))

```

Calculate the standard deviation 
```{r}
samp_dist_sd20 <-
  do(reps) * sd(~runtimeMinutes, data = sample_n(filter(d3, Decade=="20"), size = n, replace = FALSE))

samp_dist_sd30 <-
  do(reps) * sd(~runtimeMinutes, data = sample_n(filter(d3, Decade=="30"), size = n, replace = FALSE))

samp_dist_sd40 <-
  do(reps) * sd(~runtimeMinutes, data = sample_n(filter(d3, Decade=="40"), size = n, replace = FALSE))

samp_dist_sd50 <-
  do(reps) * sd(~runtimeMinutes, data = sample_n(filter(d3, Decade=="50"), size = n, replace = FALSE))

samp_dist_sd60 <-
  do(reps) * sd(~runtimeMinutes, data = sample_n(filter(d3, Decade=="60"), size = n, replace = FALSE))

samp_dist_sd70 <-
  do(reps) * sd(~runtimeMinutes, data = sample_n(filter(d3, Decade=="70"), size = n, replace = FALSE))
```

Calculate the MEAN and the STANDARD DEVIATION of the sampling distribution of sample means for each decade

```{r}
# Means
mean20 <- mean(samp_dist_mean20$mean)

mean30 <- mean(samp_dist_mean30$mean)

mean40 <- mean(samp_dist_mean40$mean)

mean50 <- mean(samp_dist_mean50$mean)

mean60 <- mean(samp_dist_mean60$mean)

mean70 <- mean(samp_dist_mean70$mean)

# Standard Deviations
sd20 <- sd(samp_dist_mean20$mean)

sd30 <- sd(samp_dist_mean30$mean)

sd40 <- sd(samp_dist_mean40$mean)

sd50 <- sd(samp_dist_mean50$mean)

sd60 <- sd(samp_dist_mean60$mean)

sd70 <- sd(samp_dist_mean70$mean)

```


Plot a histogram of the sampling distribution. What shape does it have?

```{r}
hist(samp_dist_mean20$mean)

hist(samp_dist_mean30$mean)

hist(samp_dist_mean40$mean)

hist(samp_dist_mean50$mean)

hist(samp_dist_mean60$mean)

hist(samp_dist_mean70$mean)
```
- The distributions of each decade look relatively normally distributed. The 20s decade is slightly left skewed. 

Finally, compare the standard error in runtimeMinutes for samples of size 100 from each decade [1] as estimated from your first sample of 100 movies, [2] as calculated from the known population standard deviations for each decade, and [3] as estimated from the sampling distribution of sample means.

```{r}
# 1st- Find the SE 
SE20 <- samp_dist_sd20 / sqrt(length(100))
SE30 <- samp_dist_sd30 / sqrt(length(100))
SE40 <- samp_dist_sd40 / sqrt(length(100))
SE50 <- samp_dist_sd50 / sqrt(length(100))
SE60 <- samp_dist_sd60 / sqrt(length(100))
SE70 <- samp_dist_sd70 / sqrt(length(100))

```



#                                    Challenge 2
#                                    -----------
Based on previous knowldge, she believes that the mean number of foragers that will return in that time is 12, roughly one every 6 minutes. Let  X represent the appropriate Poisson random variable of the number of arriving bees each monitoring session.
```{r}
l <-  12
p1 <-
  plotDist(
    "pois",
    lambda = l,
    main = paste0("Poisson Distribution\nwith lambda=", l),
    xlab = "x",
    ylab = "Pr(X=x)"
  )
  
p1
```



What is the probability that she will see 9 or fewer bees arrive during any given session?
```{r}
ppois(q = 9, lambda = 12, lower.tail = TRUE) 
```
24%

What is the probability that she will see no bees arrive in a session?
```{r}
ppois(q = 0, lambda = 12, lower.tail = TRUE) 
```
0% chance


What is the probability that she will see exactly 5 bees arrive in a session?
```{r}
dpois(x = 5, lambda = 12)
```
1%

What is the probability that she will see more than 18 bees arrive in a session?

```{r}
ppois(q = 18, lambda = 12, lower.tail = FALSE) 
```
4%

Plot the relevant Poisson mass function over the values in range 0 ≤  
x  ≤ 24.

```{r}
success <- c(0:24)
plot1 <- plot(success, dpois(success, lambda = 12), type='h', ylab="Probability", xlab="Number of Successes")
```

Using the rpois() function, simulate 1460 results from this distribution (i.e., 4 full years of morning monitoring sessions).

```{r}
r1 <- rpois(n=1460, lambda = 12)
```


Plot the simulated results using the histogram() function from the {mosaic} package and use xlim() to set the horizontal limits to be from 0 to 24.

```{r}
histogram(r1,xlim(0-24))
```

How do your simulated results compare to the shape of the probability mass function you plotted above?
 - My simulated results look very similar to the PMF plot, but may be a bit skewed to the right.  

                                        Challenge 3
                                        ------------
                                        
                                        
                                        
```{r}
r <- "https://raw.githubusercontent.com/difiore/ada-2021-datasets/master/zombies.csv"
z <- read.csv(r)
z
```
                                        
Calculate the population mean and standard deviation for each quantitative random variable (height, weight, age, number of zombies killed, and years of education).

```{r}
# Height mean and SD
reps <- 500
n <-  10000
samp_dist_mean_height <-
  do(reps) * mean(~height, data = sample_n(z, size = n, replace = TRUE))
samp_dist_mean_height
height_pop_mean<- mean(~mean, data = samp_dist_mean_height)
height_pop_mean
sdpop(z$height)
```
The height population mean is 67.2
The standard deviation is 4.3


```{r}
# Weight mean and SD
reps <- 500
n <-  10000
samp_dist_mean_weight <-
  do(reps) * mean(~weight, data = sample_n(z, size = n, replace = TRUE))
samp_dist_mean_weight
weight_pop_mean<- mean(~mean, data = samp_dist_mean_weight)
weight_pop_mean
sdpop(z$weight)
```

The weight population mean is 143.9
The standard deviation is 18.4



```{r}
# Age mean and SD
reps <- 500
n <-  10000
samp_dist_mean_age <-
  do(reps) * mean(~age, data = sample_n(z, size = n, replace = TRUE))
samp_dist_mean_age
age_pop_mean<- mean(~mean, data = samp_dist_mean_age)
age_pop_mean
sdpop(z$age)
```
The age population mean is 20.0
The standard deviation is 3.0

```{r}
# Zombies killed mean and SD
reps <- 500
n <-  10000
samp_dist_mean_zombie <-
  do(reps) * mean(~zombies_killed, data = sample_n(z, size = n, replace = TRUE))
samp_dist_mean_zombie
zombie_pop_mean<- mean(~mean, data = samp_dist_mean_zombie)
zombie_pop_mean
sdpop(z$zombies_killed)
```
The number of zombies killed popualtion mean is 3.0
The standard deviation is 1.7

```{r}
# Years of education mean and SD
reps <- 500
n <-  10000
samp_dist_mean_educ <-
  do(reps) * mean(~years_of_education, data = sample_n(z, size = n, replace = TRUE))
samp_dist_mean_educ
educ_pop_mean<- mean(~mean, data = samp_dist_mean_educ)
educ_pop_mean
sdpop(z$years_of_education)
```
The years of education population mean is 3.0
The standard deviation is 1.7




Use {ggplot} and make boxplots of each of these variables by gender.

```{r}
ggplot(z,aes(x=gender, y=height))+geom_boxplot()
```

```{r}
ggplot(z,aes(x=gender, y=weight))+geom_boxplot()
```
```{r}
ggplot(z,aes(x=gender, y=age))+geom_boxplot()
```
```{r}
ggplot(z,aes(x=gender, y=zombies_killed))+geom_boxplot()
```
```{r}
ggplot(z,aes(x=gender, y=years_of_education))+geom_boxplot()
```

Use {ggplot2} and make scatterplots of height and weight in relation to age, using different colored points for males versus females. 

```{r}
ggplot(z,aes(x=height, y=age, color=gender))+geom_point()
```

```{r}
ggplot(z,aes(x=weight, y=age, color=gender))+geom_point()
```
 Do these variables seem to be related? In what way?
 - Both height and weight are directly correlated with age for both male and female.

Using histograms and Q-Q plots, check whether the quantitative variables seem to be drawn from a normal distribution.

```{r}
histogram(z$height)
qqnorm(z$height)
```
```{r}
histogram(z$weight)
qqnorm(z$weight)
```

```{r}
histogram(z$age)
qqnorm(z$age)
```

```{r}
histogram(z$zombies_killed)
qqnorm(z$zombies_killed)
```
```{r}
histogram(z$years_of_education)
qqnorm(z$years_of_education)
```

Height, weight, and age are all normally distributed. Number of zombies killed looks like a gamma distribution. Years of education is not normally distributed either. Maybe its a gamma distribution?


Now use the sample_n() function from {dplyr} to sample ONE subset of 50 zombie apocalypse survivors (without replacement)
```{r}
s1 <- sample_n(tbl=z,50, replace=FALSE)
s1
```

Calculate the mean and sample standard deviation for each variable
```{r}
m.h <- mean(s1$height)
sd.h <- sd(s1$height)
m.w <- mean(s1$weight)
sd.w <- sd(s1$weight)
m.z <- mean(s1$zombies_killed)
sd.z <- sd (s1$zombies_killed)
m.e <- mean(s1$years_of_education)
sd.e <- sd(s1$years_of_education)
m.a <- mean(s1$age)
sd.a <- sd(s1$age)
```
Mean height = 68.7
SD of height = 4.8
mean weight = 149.4
Sd of weight = 19.6
Mean zombies killed = 2.7
Sd of zombies killed = 1.6
mean years of education = 3.7
SD of years of education = 1.6
Mean age = 20.6
SD of age = 3.1

Estimate the standard error for each variable based on this sample

```{r}
sigma <- 4.8
(se.height <- sigma / sqrt(length(s1))) 
sigma <- 19.6
(se.weight <- sigma / sqrt(length(s1))) 
sigma <- 1.6
(se.killed <- sigma / sqrt(length(s1))) 
sigma <- 1.6
(se.education <- sigma / sqrt(length(s1))) 
sigma <- 3.1
(se.age <- sigma / sqrt(length(s1))) 
```
SE height = 1.3
SE weight = 6.2
SE zombies killed = 0.5
SE years of education = 0.5
SE age = 1.0


Construct a 95% confidence interval for each mean
```{r}
# CI for Height
percent_ci <- 95
alpha <- 1-percent_ci / 100
lower <- m.h + qnorm(alpha /2)*se.height
lower
upper <- m.h + qnorm(1 - alpha / 2) * se.height
upper
```

CI for Height = 66.2 -71.2

```{r}
# CI for Weight
percent_ci <- 95
alpha <- 1-percent_ci / 100
lower <- m.w + qnorm(alpha /2)*se.weight
lower
upper <- m.w + qnorm(1 - alpha / 2) * se.weight
upper
```
CI for Weight = 137.3 - 161.6

```{r}
# CI for Years of Education
percent_ci <- 95
alpha <- 1-percent_ci / 100
lower <- m.e + qnorm(alpha /2)*se.education
lower
upper <- m.e + qnorm(1 - alpha / 2) * se.education
upper
```
CI for years of education = 2.7 - 4.7 

```{r}
# CI for number of zombies killed
percent_ci <- 95
alpha <- 1-percent_ci / 100
lower <- m.z + qnorm(alpha /2)*se.killed
lower
upper <- m.z + qnorm(1 - alpha / 2) * se.killed
upper
```
CI for number of zombies killed = 1.7 - 3.7

```{r}
# CI for age
percent_ci <- 95
alpha <- 1-percent_ci / 100
lower <- m.a + qnorm(alpha /2)*se.age
lower
upper <- m.a + qnorm(1 - alpha / 2) * se.age
upper
```
CI for age = 18.7 - 22.5



Then draw another 99 random samples of 50 zombie apocalypse
What are the means and standard deviations of the sampling distribution for each variable?


```{r}
k <- 99
n <- 50
s <- list()

# Height mean
means_height=numeric(k)
for(i in 1:k){
  s[[i]] <- sample_n(z, size=n, replace=FALSE)}
for(i in 1:k){
  means_height[i]=mean(s[[i]]$height)
}
means_height

# Height SD
SDs_height=numeric(k)
for(i in 1:k){
  s[[i]] <- sample_n(z, size=n, replace=FALSE)}
for(i in 1:k){
  SDs_height[i]=sd(s[[i]]$height)
}
SDs_height

# Weight Mean
means_weight=numeric(k)
for(i in 1:k){
  s[[i]] <- sample_n(z, size=n, replace=FALSE)}
for(i in 1:k){
  means_weight[i]=mean(s[[i]]$weight)
}
# Weight SD
SDs_weight=numeric(k)
for(i in 1:k){
  s[[i]] <- sample_n(z, size=n, replace=FALSE)}
for(i in 1:k){
  SDs_weight[i]=sd(s[[i]]$weight)
}

# Age mean
means_age=numeric(k)
for(i in 1:k){
  s[[i]] <- sample_n(z, size=n, replace=FALSE)}
for(i in 1:k){
  means_age[i]=mean(s[[i]]$age)
}
# Age SD
SDs_age=numeric(k)
for(i in 1:k){
  s[[i]] <- sample_n(z, size=n, replace=FALSE)}
for(i in 1:k){
  SDs_age[i]=sd(s[[i]]$age)
}

# Zombies Killed mean
means_zombiesK=numeric(k)
for(i in 1:k){
  s[[i]] <- sample_n(z, size=n, replace=FALSE)}
for(i in 1:k){
  means_zombiesK[i]=mean(s[[i]]$zombies_killed)
}
# Zombies Killed SD
SDs_zombiesK=numeric(k)
for(i in 1:k){
  s[[i]] <- sample_n(z, size=n, replace=FALSE)}
for(i in 1:k){
  SDs_zombiesK[i]=sd(s[[i]]$zombies_killed)
}

# Years of Education Mean

means_edu=numeric(k)
for(i in 1:k){
  s[[i]] <- sample_n(z, size=n, replace=FALSE)}
for(i in 1:k){
  means_edu[i]=mean(s[[i]]$years_of_education)
}
# Years of Education SD
SDs_edu=numeric(k)
for(i in 1:k){
  s[[i]] <- sample_n(z, size=n, replace=FALSE)}
for(i in 1:k){
  SDs_edu[i]=sd(s[[i]]$years_of_education)
}

```


Finally, construct an 95% confidence interval for each mean directly from the sampling distribution of sample means using the central 95% that distribution (i.e., by setting the lower and upper CI bounds to 2.5% and 97.5% of the way through that distribution).


```{r}
# The standard deviation of the sampling distribution is the standard error

ci_height <- means_height + c(-1, 1) * qnorm(1 - alpha / 2) * SDs_height

ci_weight <- means_weight + c(-1, 1) * qnorm(1 - alpha / 2) * SDs_weight


ci_edu <- means_edu + c(-1, 1) * qnorm(1 - alpha / 2) * SDs_edu

ci_zombiesK <- means_zombiesK + c(-1, 1) * qnorm(1 - alpha / 2) * SDs_zombiesK

ci_age <- means_age + c(-1, 1) * qnorm(1 - alpha / 2) * SDs_age
```



How do the standard deviations of the sampling distribution for each variable compare to the standard errors estimated from your first sample of size 50? What do sampling distributions for each variable mean look like? Are they normally distributed? What about for those variables that you concluded were not originally drawn from a normal distribution? 
```{r}
#Big Sample HEIGHT
hist(SDs_height) # I would consider this pretty much normally distributed
mean(SD_height)
# the mean of the big sample SDs = 4.24

#Small Sample
sd.h # the small sample SD = 4.04




# Big Sample WEIGHT
hist(SDs_weight) # The curve has a normal shape 
mean(SDs_weight) # the mean of the SDs is 18.3

sd.w # The SD of the small sample is 17


# Big Sample AGE
hist(SDs_age)
mean(SDs_age) # 2.9

sd.a # 2.5


#Big Sample YEARS OF EDUCATION
hist(SDs_edu)
mean(SDs_edu) # 1.7

sd.e # 1.7

# Big Sample ZOMBIES KILLED
hist(SDs_zombiesK)
mean(SDs_zombiesK) #1.8

sd.z # 1.8



# All of the distributions look normal.
# The SDs of the large sample are very close to the SDs of the smaller sample. If anything, the SDs of the big sample are a bit larger than those of the small sample. 
```

How do the two 95% CIs you estimated compare to one another (i.e., the CI based on one sample and its estimated standard deviation versus the CI based on simulation)?

```{r}
mean(ci_height) #67.5
# The mean of the CIs for height from the big sample falls right within the CI range of the small sample

mean(ci_weight) # 143.8
#The mean of the CIs for weight from the big sample falls right within the CI range of the small sample

mean(ci_age)# 20.0
#The mean of the CIs for age from the big sample falls right within the CI range of the small sample

mean(ci_zombiesK) # 2.9
#The mean of the CIs for zombies killed from the big sample falls right within the CI range of the small sample

mean(ci_edu) #3.0
#The mean of the CIs for years of education from the big sample falls right within the CI range of the small sample
```


